\Huge \textbf{Hybrid rendering}\\ \huge using Radeon Rays

### 2018-11-03 Paper/Paper_BillietStijn_RadeonRays.tex
Fixed a load of typos (many more to come i'm afraid).

\subsubsection{Definition}
Rasterization is an algorithm designed to display three-dimensional objects on a two-dimensional screen. It's very fast seeing this algorithm is parallel in nature and can thus make full use of the GPU. The results of rasterization have gotten very good over the years, very convincing at relatively low compute performance. However there are certain limitations to this algorithm, seeing it's main goal is to get graphics in screen space.\newline\par

@@

\subsubsection{Rasterization stage}
The fourth stage is rasterization, this is where the algorithm gets it's name. The rasterizer is in charge of a not only rasterization, but also homogeneous divide, clipping, culling and viewport transform. \newline\par

\begin{minipage}{0.5\textwidth}
	\centering
	\includegraphics[width=\textwidth]{culling.png}
\end{minipage} \hfill
\begin{minipage}{0.5\textwidth}
	The first task of the rasterizer is primitive culling. The shapes that made it to the rasterizer are checked based on their normals. The normal of a primitive/vertex tells us if the face is facing away from us or not. This simple operation can be demonstrated using the dotproduct, however this does get handled by the specifications (e.g. DirectX/OpenGL) rasterizer. In it's essence the calculation will translate to if the dotproduct is negative then the face is facing away, if it's positive then it's facing toward us.
\end{minipage}

@@

A great thing about this control of memory is that it empowers the developer to make use of interoperation contexts. This can be used to achieve zero-copy architectures where memory objects can be shared by multiple compatible api's. A prime example of this is an OpenCL/OpenGL interop context. Such a context allows the graphics programmers to send screen buffers, textures and openGL objects to the kernels for further processing without the need to copy them, further enabling them to use a kernel in the pipeline everywhere they want.\newline\par

While this is great, there are still some limitations that have to be set in place to enable this kind of behaviour. For one the target device needs to support this sharegroup behaviour and for seconds shared memory can cause issues seeing that both contexts don't know about each others existance. All of this can be alliviated by setting blocking calls (e.g. GPU fences) in place, with these acting as a sort of failsafe that makes sure the object is done reading from/ writing to by the target context. This can account for slowdowns seeing that the CPU is needed to help the 2 parties syncronize, luckily for the programmer there is hope that said GPU supports the ARB\_cl\_event extension.3 The event extension allows the two api's to communicate through event management (e.g. wait for event to begin) ensuring that they don't need to wait a few ticks of CPU-time to synchronize.

@@

	There are numerous libraries out there that do just that, libraries like GLEW, GL3W, Glad, Glee and many more. All libraries that exist out there all deliver on the same promise, meaning that the choice of library really boils down to personal preference.\newline\par
	\item{\textbf{GLM}}\newline\par
	When it comes to math libraries that plug in nicely with an OpenGL renderer there's really only one good option namely GLM. OpenGL Mathematics (GLM) is a header only C++ mathematics library for graphics software based on the OpenGL Shading Language (GLSL). GLM exposes a well defined interface for everything from general calculus to quaternion rotations (e.g. vector operations, matrix calculations).
\end{itemize}
\subsubsection{Window creation}
A window allows the rendering specification to offload its framebuffer to the viewport, it also acts as a layer between the operating system and the game when it comes to input etc. The creation of a window is entirely dependent on the operating system it's creating the window for. For example when creating a window on linux the X window system should be addressed, whereas on Windows you would need to hook on to the WinMain function and use the WndProc callback to fetch any input. Therefore it is much smarter to use a library that hides all this operating system agnostic stuff in an abstraction layer.\newline\par
SDL, SFML and GLFW are all low level libraries that handle multimedia indirection. They all deliver the same features in some shape or form. The choice really boils down to maturity of the library, documentation, usefulness of its features and frequency of the updates (if any).
\subsubsection{Texture loading}
A texture is a 2D (could be 3D or 1D) image packed with color data that is used to feed into the graphics pipeline to texturize geometry. Textures can be stored in a dozen of file formats, each with their own structure to store the pixel data. Textures are stored in a binary format and can therefore be parsed by a binary reader. The problem however with doing textureloading is that every format is different so the programmer would be in charge of writing a dozen different loaders to support the most basic of texture types. Seeing that this is rather cumbersome most people would opt for a library that does this for them. Again there are quite a lot of libraries out there that do this. \newline\par

@@

The third rendering pipeline, although the most performant will not be covered with explicit detail. The reason that this technique is not that applicable has primarily to do with the fact that we have no way of doing per world space lighting effects. Forward+ is a rendering technique that improves upon regular forward rendering. Forward+ first determines which lights are overlapping which areas of the screen (screen space). During the shading phase that list is then consulted to only account for the lights that are potentially overlapping the current pixel fragment. The forward+ technique consists primarily of three simple passes. In the light culling pass, each light in the scene is sorted into screen space tiles. In the opaque pass, the light list generated from the light culling pass is used to compute the lighting for opaque geometry. The transparent pass is similar to the opaque pass in the sense that the light list is again consulted when shading the transparent pixel fragments.

@@

With the geometry data now in the g-buffers, that data can now be used to construct the rays. The ray construction can be sped up by doing it on the GPU, seeing that this operation is highly parallel in nature. Essentially what it boils down to is to make sure that meshes have been synchronized with the raytracer, meaning that the raytracer has all the right meshes in it's scene to be able to fetch all occlusion/intersections. Then quite simply sample all the position entries and normal entries (for the entire screen) and send out a ray for every pixel, this will in turn correlate to the raytracer api seeing those position entries should correspond with the meshes in the scene.\newline\par

>>>>

\subsubsection{Definition}
Rasterization is an algorithm designed to display three-dimensional objects on a two-dimensional screen. It's very fast seeing this algorithm is parallel in nature and can thus make full use of the GPU. The results of rasterization have gotten very good over the years, very convincing at relatively low compute performance. However there are certain limitations to this algorithm, seeing its main goal is to get graphics in screen space.\newline\par

@@

\subsubsection{Rasterization stage}
The fourth stage is rasterization, this is where the algorithm gets its name. The rasterizer is in charge of a not only rasterization, but also homogeneous divide, clipping, culling and viewport transform. \newline\par

\begin{minipage}{0.5\textwidth}
	\centering
	\includegraphics[width=\textwidth]{culling.png}
\end{minipage} \hfill
\begin{minipage}{0.5\textwidth}
	The first task of the rasterizer is primitive culling. The shapes that made it to the rasterizer are checked based on their normals. The normal of a primitive/vertex tells us if the face is facing away from us or not. This simple operation can be demonstrated using the dotproduct, however this does get handled by the specifications (e.g. DirectX/OpenGL) rasterizer. In its essence the calculation will translate to if the dotproduct is negative then the face is facing away, if it's positive then it's facing toward us.
\end{minipage}

@@

A great thing about this control over memory is that it empowers the developer to make use of interoperation contexts. This can be used to achieve zero-copy architectures where memory objects can be shared by multiple compatible api's. A prime example of this is an OpenCL/OpenGL interop context. Such a context allows the graphics programmers to send screen buffers, textures and openGL objects to the kernels for further processing without the need to copy them, further enabling them to use a kernel in the pipeline everywhere they want.\newline\par

While this is great, there are still some limitations that have to be set in place to enable this kind of behaviour. For one the target device needs to support this sharegroup behaviour and for seconds shared memory can cause issues seeing that both contexts don't know about each others existance. All of this can be alliviated by setting blocking calls (e.g. GPU fences) in place, with these acting as a sort of failsafe that makes sure the object is done reading from/ writing to by the target context. This can account for slowdowns seeing that the CPU is needed to help the 2 parties synchronize, luckily for the programmer there is hope that said GPU supports the ARB\_cl\_event extension.3 The event extension allows the two api's to communicate through event management (e.g. wait for event to begin) ensuring that they don't need to wait a few ticks of CPU-time to synchronize.

@@

	There are numerous libraries out there that do just that, libraries like GLEW, GL3W, Glad, Glee and many more. All libraries that exist out there all deliver on the same promise, meaning that the choice of library really boils down to personal preference.\newline\par
	\item{\textbf{GLM}}\newline\par
	When it comes to math libraries that plug in nicely with an OpenGL renderer there's really only one good option namely GLM. OpenGL Mathematics (GLM) is a header only C++ mathematics library for graphics software based on the OpenGL Shading Language (GLSL). \newline\par
	GLM exposes a well defined interface for everything from calculus to quaternion rotations (e.g. vector operations, matrix calculations).
\end{itemize}
\subsubsection{Window creation}
A window allows the rendering specification to offload its framebuffer to the viewport, it also acts as a layer between the operating system and the game when it comes to input etc. The creation of a window is entirely dependent on the operating system it's creating the window for. For example when creating a window on linux the X window system should be addressed, whereas on Windows you would need to hook on to the WinMain function and use the WndProc callback to fetch any input. Therefore it is much smarter to use a library that hides all this operating system agnostic stuff in an abstraction layer.\newline\par
SDL, SFML and GLFW are all low level libraries that handle multimedia indirection. They all deliver the same features in some shape or form. The choice really boils down to maturity of the library, documentation, usefulness of its features and frequency of the updates (if any).
\subsubsection{Texture loading}
A texture is a 2D (could be 3D or 1D) image packed with color data that is used to feed into the graphics pipeline to texturize geometry. Textures can be stored in a dozen of file formats, each with their own structure to store the pixel data. Textures are stored in a binary format and can therefore be parsed by a binary reader. The problem however with doing textureloading is that every format is different so the programmer would be in charge of writing a dozen different loaders to support the most basic of texture types. Seeing that this is rather cumbersome most people would opt for a library that does this for them. Again there are quite a lot of libraries out there that do this. \newline\par

@@

The third rendering pipeline, although the most performant will not be covered with explicit detail. The reason that this technique is not that applicable has primarily to do with the fact that we have no way of doing world space lighting effects. Forward+ is a rendering technique that improves upon regular forward rendering. Forward+ first determines which lights are overlapping which areas of the screen (screen space). During the shading phase that list is then consulted to only account for the lights that are potentially overlapping the current pixel fragment. The forward+ technique consists primarily of three simple passes. In the light culling pass, each light in the scene is sorted into screen space tiles. In the opaque pass, the light list generated from the light culling pass is used to compute the lighting for opaque geometry. The transparent pass is similar to the opaque pass in the sense that the light list is again consulted when shading the transparent pixel fragments.

@@

With the geometry data now in the g-buffers, that data can now be used to construct the rays. The ray construction can be sped up by doing it on the GPU, seeing that this operation is highly parallel in nature. Essentially what it boils down to is to make sure that meshes have been synchronized with the raytracer, meaning that the raytracer has all the right meshes in its scene to be able to fetch all occlusion/intersections. Then quite simply sample all the position entries and normal entries (for the entire screen) and send out a ray for every pixel, this will in turn correlate to the raytracer api seeing those position entries should correspond with the meshes in the scene.\newline\par
